Ear canal dynamic motion piezoelectric energy harvester using a bistable resonator cycled by coupled hydraulic switches made of collapsed flexible tubes.

### 2022-10-17 article_system.tex
update delete blades design
update

% for cutting words
\usepackage[utf8]{inputenc}
% a font close to that chosen by Elsevier
\usepackage[bitstream-charter]{mathdesign}
\usepackage{mathtools, cuted}
% for accentuated characters
%\usepackage[T1]{fontenc}
\geometry{top=2cm,left=1.5cm,bottom=2cm,right=1.5cm}
\usepackage{multicol}
% For treating ps:
\usepackage{psfrag}
% For subfigures
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{placeins}
% For tables
\usepackage{array}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{colortbl}
\usepackage{tabulary}
\usepackage{etoolbox}
% For citations
\usepackage[numbers]{natbib}
%\biboptions{longnamesfirst,angle,semicolon}
%\citetstyle{nature}
% For color
\usepackage{color}
\usepackage{xcolor}
% For spacing
\usepackage{setspace}
%\doublespacing
\singlespacing
%\onehalfspacing
% For maths
\usepackage{amsmath, amsfonts}
\usepackage[squaren,ashgrey]{SIunits}
% For Line numbering
\usepackage[pagewise,modulo]{lineno}
%\linenumbers
% For special characters
\usepackage{pifont}
%
\usepackage[pagebackref,colorlinks=True]{hyperref}
\usepackage{wasysym}
\usepackage{color}
\usepackage{cleveref}
\crefformat{figure}{(Fig.~#2#1#3)}
\usepackage{hyperref}
\usepackage[hyperpageref]{backref}
\usepackage{lipsum}
\usepackage{booktabs} % To thicken table lines
% \usepackage{widetext}
\usepackage{float}
\usepackage{empheq}
\usepackage{textcomp}
\usepackage{multicol}

@@

\address[symme]{Laboratoire SYMME - Université Savoie Mont Blanc, 7 Chemin de Bellevue, 74940, Annecy}
\address[critias]{Laboratoire Critias - École de Technologie Supérieure, 1100 Rue Notre-Dame Ouest, Montréal, QC, H3C 1K3}

@@

\author[critias]{Michel DEMUYNCK}
\author[critias]{Aidin DELNAVAZ}
\author[critias]{Jérémy VOIX}

@@

		HC                       & Hydraulic Cylinder                                      \\

@@

@@

		$a$                      & HV bending lever arm                                    \\

@@

		$Cf$ [Pam$^3$/s]         & Pressure loss coefficient                               \\
		$Cf_0$ [Pam$^3$/s]       & Opened HV pressure loss coefficient                     \\
		$Cf_c$ [Pam$^3$/s]       & Closed HV pressure loss coefficient                     \\

@@

		$K_{HV}$ [Nm/rad]        & HV stiffness                                            \\

@@

		$K_{\varphi}$ [Nm/rad]   & Single hinge stiffness                                  \\

@@

		$Q$                      & BR quality factor                                       \\

@@

@@

		$\mu$ [Ns/m]             & Air dynamic viscosity                                   \\
		$\mu_{hc}$ [Ns/m]        & HC viscous loss coefficient                             \\
		$\nu$ [Ns/m]             & Dry friction loss coefficient                           \\

@@

Based on the analysis of the literature, we propose a new solution optimizing the energy conversion efficiency to maximize the harvested energy from the ear canal dynamic motion. The next section states on the hydraulic energy source characterization and on the harvesting strategy we adopted. Section \ref{sec:HARVESTER PRESENTATION AND OPERATION PRINCIPLE} presents the harvester and describes its operation principle. Section \ref{sec:SYSTEM MODELING AND SIMULATIONS} shows the global system  modeling and the numerical simulation on its behavior. Then section \ref{sec:EXPERIMENTAL CHARACTERIZATIONS} then exposes the experimental validation performed on the harvester critical components and section \ref{sec:MODEL RECALIBRATION WITH EXPERIMENTAL DATA} presents the experimentally adjusted global harvester model. Section \ref{sec:ANALYZE AND DISCUSSION} discusses this work major contributions and the harvester potential. Finally, a conclusion is stated in the last section.

@@

By isolating the BR mass and using the Newtons 2$^{nd}$ law, we can express its dynamic equilibrium. The projection on the $\vec{x}$ axis of the resulting expression is given by the equation \ref{eq:OB-GPA}. The BR mechanical equilibrium has a mass inertial term, a non-linear stiffness depending on $x_m$, the stiffness of the hinge $K_{\varphi}$ and viscous losses. The first term introduced by the HV is related to its rotational stiffness $K_{HV}$ and the second term is related to the friction losses at the contact point $T$ (fig. \ref{fig:HV_actuation_detail}). A dry friction model is chosen to describe the dissipation effect of the contact. The rubbing efforts depend on a dry friction coefficient $\nu$, related to the nature of the material that are rubbing, and the sign of $\dot{x}_m$. The pressure $p_{hc,a}$ induces the HC hydraulic force and the coefficient $\mu_{hc}$ models the viscous friction at the sealing gasket. The hydro-mechanical coupling will be detailed later.

The electrical power generated by the APG is dissipated in a load resistor $R_l$ calculated with an adaptive impedance matching expressed in the equation \ref{eq:R_ch_adaptive_imp} \cite{Liu2013}. 
\begin{equation}
	R_l = \dfrac{1}{C_p w_0}
	\label{eq:R_ch_adaptive_imp}
\end{equation}
Thus, the electrical equation of the APG is :
\begin{equation}
	\dfrac{U_p}{R_l} = 
	-\alpha\dfrac{d}{dt}\biggl(2\sqrt{{l_0}^2-x_m^2}\biggr)
	- C_p\dot{U_p}
	= \frac{2\alpha x_m\dot{x}_m}{\sqrt{L^2+x_m^2}} - C_p\dot{U_p}
\end{equation}
Consequently, the generated power $P_e$ expression is :
\begin{equation}
	P_e = \frac{{U_p}^2}{R_l} 
	\label{eq:P_e}
\end{equation} 
Equation \ref{eq:theta_fonction_de_xm} defines the implicit relation between $\theta$, $x_m$ and the geometrical parameters of the setup. Also, the HV stiffness $K_{HV}$ will depend on its bending angle $\theta$ and the latter will geometrically depend on the BR mass position $x_m$, as showed in figure \ref{fig:HV_actuation_detail}. 
%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htb]
\begin{equation}
 m \ddot{x}_m =-2K(\sqrt{x_m^2+L^2}-l_0)\tan(\varphi) -\mu \dot{x}_m -\frac{4K_{\varphi}\varphi}{L}
				\ -2\alpha U_p \tan(\varphi)
				\ -\dfrac{K_{HV}(x_m)\theta(x_m)}{a} - \nu \text{sign}(\dot{x}_m)
				\  - p_{hc,a}\ S_{hc} - \mu_{c}\ \dot{x}_m
\label{eq:OB-GPA}
\end{equation}
\end{figure*}
% \FloatBarrier
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%
% \begin{figure*}
% \begin{equation}
% % \begin{small}
% 	\begin{dcases}
% \ddot{x}_m = - \frac{2K{x_0}^2}{mL^2}\Biggl[\biggl(\frac{x_m^2}{{x_0}^2} -1\biggr )+ \dfrac{8K_{\varphi}}{L^2}\Biggr]x_m 
% 			 - \frac{K_{HV}(x_m)\theta(x_m)}{ma}
% 			 - \frac{\mu+\mu_{hc}}{m}\dot{x}_m  - \nu \text{sign}(\dot{x}_m)
% 			 - \frac{2\alpha U_p\ x_m}{mL}
% 			 - p_{hc,a}\ S_{hc} \\
% x_m = \dfrac{D_r(1-\cos(\theta))+2a\sin(\theta)}{2\cos(\theta)}\text{,}
% 				~~ \theta\leq\arctan\biggl(\dfrac{2a}{D_r}\biggr)\\
% \dfrac{U_p}{R_l} = \frac{2\alpha x_m \dot{x}_m}{L} - C_p\dot{U}_p 
% 	\end{dcases}
% 	\label{eq:OB+GPA+VH+piston}	
% % \end{small}
% \end{equation}
% \end{figure*}
% % \FloatBarrier
% %%%%%%%%%%%%%%%%%%%%%%%
    %///////////////////////////////////////////// 
	\subsection{Modeling of the hydraulic circuit}	
	\label{subsec:The hydraulic circuit}
    %/////////////////////////////////////////////

@@

The no leakage assumption imposes that all the fluid exiting the earplug is contained in the two hydraulic branches (eq. \ref{eq:conservation_masse}).
\begin{equation}
	q_{ear} = \dot{\Delta V_{ear}} = q_a + q_i
	\label{eq:conservation_masse}
\end{equation}
Moreover, the fluid entering the HC necessarily results on a piston displacement (eqs. \ref{eq:conservation_masse_ouvert},\ref{eq:conservation_masse_ferme}).
\begin{align}
	\text{Active cylinder ~}& \left\{~
	q_a = S_{hc} \dot{x}_m
	\right.
	\label{eq:conservation_masse_ouvert}\\
	\text{Inactive cylinder ~}& \left\{~
	q_i = S_{hc} \dot{x}_{ic}
	\right.
	\label{eq:conservation_masse_ferme}
\end{align}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{NUMERICAL MODEL AND SIMULATIONS}
\label{NUMERICAL MODEL AND SIMULATIONS}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
    %///////////////////////////////////////////// 
	\subsection{The harvester parameters setting}	
	\label{subsec:The harvester setting}
    %/////////////////////////////////////////////
The energy system combines many variable parameters. In order to determine them, it is first necessary to identify the fixed technological elements: 
\begin{itemize}
	\item The transducer is a APA50XS piezoelectric actuator, from Cedrat Technologies, exploited as a generator \cite{CEDRATTECHNOLOGIES2022}.
	\item The hydraulic actuation is ensured by the SMC MQP4-10S HCs which admit operates at 1kPa pressure \cite{SMC2022}.
	\item The usual hearing aid size is $50$x$20$x$10$mm \cite{Quattro2019}. The \mbox{$L=16mm$} parameter is chosen to be consistent with this scale. $\epsilon$ can still be adjusted with $x_0$.
\end{itemize}

@@

Figure \ref{fig:positions_Up_puissances_energies} additionally presents the different positions of the moving components, the APG voltage, the source and harvested powers and the different energies entering and exciting the system. The theoretical global efficiency of the harvester is evaluated at \mbox{$\eta_g=76$\%} with 20µJ harvested from one mastication. The energy losses are mostly located in the BR stage and depend on the quality factor $Q$ that has been set in accordance with previous work results on similar BR architectures \cite{Liu2013}. The energy source depends on the comfort pressure level $p_c$ that is set to 1kPa here, i.e. a fraction of the theoretical maximal value of $12$kPa \cite{Bouchard-Roy2020}. The harvester efficiency could lead to a theoretical maximum of $684$µJ extractable from the ear canal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={0cm 0cm 0cm 0.5cm},clip, width=\textwidth]{figures/positions+DeltaV_debits_pear.pdf}
	\caption{}
	\label{fig:positions+DeltaV_debits_pear}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={0cm 0cm 0cm 1cm},clip, width=\textwidth]{figures/positions_Up_puissances_energies.pdf}
	\caption{}
	\label{fig:positions_Up_puissances_energies}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
\begin{table}	
	\centering
	\begin{tabular}{c|c|c}
\toprule
\multicolumn{1}{c}{~~\textbf{Parameter}~~}  & \multicolumn{1}{c}{~~\textbf{Value}~~} & \multicolumn{1}{c}{~~\textbf{[Unit]}~~}  \\
\midrule
$K$					&	0.256			& [$N/\mu m$]			\\ \hline
$x_0$				&	0.49			& [$mm$] 				\\ \hline
$L$					&	16.0			& [$mm$] 				\\ \hline
$m$					&	5.88			& [$g$] 				\\ \hline
$x_{c0}$			&	0.69			& [$mm$] 				\\ \hline
$R_l$			&	6.39			& [$k\Omega$] 			\\ \hline
$\alpha$			&   0.105			& [$N/V$] 				\\ \hline
$C_p$	   			&	0.25			& [$\mu F$]				\\ \hline
$Q$ 				&	50				& [-] 					\\ \hline
$\eta_g$			& 	76				& [$\%$]		\\ 
\bottomrule
	\end{tabular}
	\caption{Electromechanical parameters}
	\label{tab:parametres électromécaniques}
\end{table}
\begin{table}
	\centering
	\begin{tabular}{c|c|c}
\toprule
\multicolumn{1}{c}{~~\textbf{Parameter}~~}  & \multicolumn{1}{c}{~~\textbf{Value}~~} & \multicolumn{1}{c}{~~\textbf{[Unit]}~~}  \\
\midrule
$D_{hc}$  	   			&	4				& [$mm$]			\\ \hline
$a_h$					&	29				& [-]				\\ \hline
$p_c$					&	1				& [$kPa$]			\\ \hline
$\Delta V{ear,m}$		& 60				& [$mm^3$]			\\ \hline
$Cf_0$					&	0.21e17			&[$Pa.s^2m^{-6}$]	\\ \hline
$Cf_c$					&	5.46e17			&[$Pa.s^2m^{-6}$]	\\
\bottomrule 
	\end{tabular}
	\caption{Hydraulic parameters}
	\label{tab:parametres_hydrauliques}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The device heart is the electromechanical converter composed of the BR implementing the APG and therefore, they have been designed with respect to space requirements. We opted for a monobloc structure with blades in order to optimize the resonator quality factor by avoiding screw assemblies. Figure \ref{fig:monobloc_face} shows the face view of the monobloc bistable resonator CAD that has been fabricated by electrical discharge machining (EDM). The 8 hinges are assured by 4 buckled blades (BB) while a vertical guide blade (GB) ensures the APG mounting and avoids any of its rotations. The efficient part composed of the blades and the central mass are $1.2$mm think. The external rigid $6$mm thick frame has been designed to ensure the fixed boundary conditions and also to allow the mounting of both the measurement instruments and the HCs. The local thickenings at the midspan of the blades are intended to facilitate the fabrication and also to maximize the BB compression stiffness that maximizes the electromechanical coupling of the harvester, as we show in the following sections. The geometrical and mechanical parameters of the BR are referenced in table \ref{tab:parametres_lames}. The following sections describe the design method that led to the dimensions of the blades that depend on our needs.

@@

Soft hinge stiffness		   & $K_{\varphi}$             & 0.006 [Nm/rad] \\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[!htbp]
\captionsetup{justification=centering}
	\includegraphics[trim={19.2cm 0cm 0cm 5cm},clip,width=.7\linewidth]{figures/monobloc_face.pdf}
	\caption{Face transparency view of the monobloc BR CAO}
\label{fig:monobloc_face}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The harvesting occurs during the free oscillation phase. The main issue is to maximize the energy conversion efficiency $\eta_{ob}$ of this stage. Richards \emph{et al.} demonstrated that for the piezoelectric transduction at the resonance frequency, it can be approximated with the equation \ref{eq:eta_ob} \cite{Richards2004}. This expression is valid for an electric extraction based on the adaptive impedance matching strategy.

@@

Where $k^2_{sys}$ is the electromechanical coupling of the harvesting system. The latter can be expressed with the electromechanical parameters of the APG and the stiffness of the respective blades (tab. \ref{tab:parametres_lames}) as follows :

@@

\begin{equation}
	k^2_{sys} = \dfrac{\alpha^2_{eq}}{\alpha^2_{eq} + (K+K_{gb})C_p}
	\label{eq:k2_sys}
\end{equation}
Where $\alpha_{eq}$ is the equivalent piezoelectric force factor (PFF) given by equation \ref{eq:alpha_eq}. It depends on the initial PFF of the APG alone and the and ratio between the equivalent stiffness $K_{eq}$ (eq. \ref{eq:K_eq}) of the structure on the $\vec{x}$ axis and the equivalent stiffness of the APG mounted on the GB.
\begin{equation}
	\alpha_{eq} = \alpha\ \dfrac{K_{eq}}{K + K_{gb}} 
	\label{eq:alpha_eq}
\end{equation}
\begin{equation}
	K_{eq} = \dfrac{K_{bb}(K + K_{gb})}{K_{bb} + K_{gb} + K}
	\label{eq:K_eq}
\end{equation}
Equations \ref{eq:eta_ob} - \ref{eq:K_eq} combined demonstrate that in order to maximize $\eta_{ob}$ we must make sure that $K_{bb} \gg K \gg K_{gb}$. If the condition is respected, then $K_{eq}$ can be assimilated to $K$.

>>>>

% for cutting words
\usepackage[utf8]{inputenc}
% a font close to that chosen by Elsevier
\usepackage[bitstream-charter]{mathdesign}
\usepackage{mathtools, cuted}
% for accentuated characters
%\usepackage[T1]{fontenc}
\geometry{top=2cm,left=1.5cm,bottom=2cm,right=1.5cm}
\usepackage{multicol}
% For treating ps:
\usepackage{psfrag}
% For subfigures
\usepackage{subcaption}
\usepackage{placeins}
% For tables
\usepackage{array}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{colortbl}
\usepackage{tabulary}
\usepackage{etoolbox}
% For citations
\usepackage[numbers]{natbib}
%\biboptions{longnamesfirst,angle,semicolon}
%\citetstyle{nature}
% For color
\usepackage{color}
\usepackage{xcolor}
% For spacing
\usepackage{setspace}
%\doublespacing
\singlespacing
%\onehalfspacing
% For maths
\usepackage{amsmath, amsfonts}
\usepackage[squaren,ashgrey]{SIunits}
% For Line numbering
\usepackage[pagewise,modulo]{lineno}
%\linenumbers
% For special characters
\usepackage{pifont}
%
\usepackage[pagebackref,colorlinks=True]{hyperref}
\usepackage{wasysym}
\usepackage{color}
\usepackage{cleveref}
\crefformat{figure}{(Fig.~#2#1#3)}
\usepackage{hyperref}
\usepackage[hyperpageref]{backref}
\usepackage{lipsum}
\usepackage{booktabs} % To thicken table lines
% \usepackage{widetext}
\usepackage{float}
\usepackage{empheq}
\usepackage{textcomp}
\usepackage{multicol}
\usepackage{graphicx}

@@

\address[symme]{Laboratoire SYMME - Université Savoie Mont Blanc, 7 Chemin de Bellevue, 74940, Annecy}
\address[critias]{Université du Québec - École de technologie supérieure, 1100 Rue Notre-Dame Ouest, Montréal, QC, H3C 1K3}

@@

\author[critias]{Michel DEMUYNCK}
\author[critias]{Aidin DELNAVAZ}
\author[critias]{Jérémie VOIX}

@@

		HC                       & Hydraulic cylinder                                      \\

@@

		PFF						 & piezoelectric force factor  \\

@@

		$a$ [m]                  & HV bending lever arm                                    \\

@@

		$Cf$ [Pa.s$^2$.m$^{-6}$]         & Pressure loss coefficient                               \\
		$Cf_0$ [Pa.s$^2$.m$^{-6}$]       & Opened HV pressure loss coefficient                     \\
		$Cf_c$ [Pa.s$^2$.m$^{-6}$]       & Closed HV pressure loss coefficient                     \\

@@

		$K_{HV}$ [N.m/rad]        & HV stiffness                                            \\

@@

		$K_{\varphi}$ [N.m/rad]   & Single hinge stiffness                                  \\

@@

		$Q$ [-]                  & BR quality factor                                       \\

@@

		$x_0$ [m]                & BR buckling height                                       \\

@@

		$\mu$ [N.s/m]             & Air dynamic viscosity                                   \\
		$\mu_{hc}$ [N.s/m]        & HC viscous loss coefficient                             \\
		$f_d$ [N.s/m]             & Dry friction loss coefficient                           \\
		$\varphi$ [rad]			 & BR hinge rotational angle							   \\

@@

Based on the analysis of the literature, we propose a new solution optimizing the energy conversion efficiency to maximize the harvested energy from the ear canal dynamic motion. The next section states on the hydraulic energy source characterization and on the harvesting strategy that we adopted. Section \ref{sec:HARVESTER PRESENTATION AND OPERATION PRINCIPLE} presents the harvester and describes its operation principle. Section \ref{sec:SYSTEM MODELING} details the global system modeling and section \ref{sec:NUMERICAL MODEL AND SIMULATIONS} presents the numerical simulation of the theoretical behavior of the harvester. Then section \ref{sec:EXPERIMENTAL CHARACTERIZATIONS OF THE ELECTROMECHANICAL CONVERTER} then exposes the experimental validation performed on the electromechanical converter stage and section \ref{sec:EXPERIMENTAL APPROACH FOR THE HV DESIGN} shows the experimental approach for the design of the hydraulic valves for the system cycling. Then, section \ref{sec:MODEL RECALIBRATION WITH EXPERIMENTAL DATA} presents the experimentally adjusted global harvester model and discusses this work major contributions and the harvester potential. Finally, a conclusion is stated in the last section.

@@

By isolating the BR mass and using the Newtons 2$^{nd}$ law, we can express its dynamic equilibrium. The projection on the $\vec{x}$ axis of the resulting expression is given by the equation \ref{eq:OB-GPA}. The BR mechanical equilibrium has a mass inertial term, a non-linear stiffness depending on $x_m$, the stiffness of the hinge $K_{\varphi}$ and viscous losses. The first term introduced by the HV is related to its rotational stiffness $K_{HV}$ and the second term is related to the friction losses at the contact point $T$ (fig. \ref{fig:HV_actuation_detail}). $K_{HV}$ and $K_{\varphi}$ both act the same way by making move the stable equilibrium positions of the BR toward $0$ as much as they increase. A dry friction model is chosen to describe the dissipation effect of the contact. The rubbing efforts depend on a dry friction coefficient $f_d$, related to the nature of the material that are rubbing, and the sign of $\dot{x}_m$. The pressure $p_{hc,a}$ induces the HC hydraulic force and the coefficient $\mu_{hc}$ models the viscous friction at the sealing gasket. The hydro-mechanical coupling will be detailed later.

The electrical power generated by the APG is dissipated in a load resistor $R_l$ calculated with an adaptive impedance matching expressed in the equation \ref{eq:R_ch_adaptive_imp} \cite{Liu2013}. 
\begin{equation}
	R_l = \dfrac{1}{C_p w_0}
	\label{eq:R_ch_adaptive_imp}
\end{equation}
Thus, the electrical equation of the APG is :
\begin{equation}
	\dfrac{U_p}{R_l} = 
	-\alpha\dfrac{d}{dt}\biggl(2\sqrt{{l_0}^2-x_m^2}\biggr)
	- C_p\dot{U_p}
	= \frac{2\alpha x_m\dot{x}_m}{\sqrt{L^2+x_m^2}} - C_p\dot{U_p}
\end{equation}
Consequently, the generated power $P_e$ expression is :
\begin{equation}
	P_e = \frac{{U_p}^2}{R_l} 
	\label{eq:P_e}
\end{equation} 
Equation \ref{eq:theta_fonction_de_xm} defines the implicit relation between $\theta$, $x_m$ and the geometrical parameters of the setup. Also, the HV stiffness $K_{HV}$ will depend on its bending angle $\theta$ and the latter will geometrically depend on the BR mass position $x_m$, as showed in figure \ref{fig:HV_actuation_detail}. 
%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htb]
\begin{equation}
 m \ddot{x}_m =-2K(\sqrt{x_m^2+L^2}-l_0)\tan(\varphi) -\mu \dot{x}_m -\frac{4K_{\varphi}\varphi}{L}
				\ -2\alpha U_p \tan(\varphi)
				\ -\dfrac{K_{HV}(x_m)\theta(x_m)}{a} - f_d \text{sign}(\dot{x}_m)
				\  - p_{hc,a}\ S_{hc} - \mu_{c}\ \dot{x}_m
\label{eq:OB-GPA}
\end{equation}
\end{figure*}
% \FloatBarrier
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%
% %%%%%%%%%%%%%%%%%%%%%%%
% \begin{figure*}
% \begin{equation}
% % \begin{small}
% 	\begin{dcases}
% \ddot{x}_m = - \frac{2K{x_0}^2}{mL^2}\Biggl[\biggl(\frac{x_m^2}{{x_0}^2} -1\biggr )+ \dfrac{8K_{\varphi}}{L^2}\Biggr]x_m 
% 			 - \frac{K_{HV}(x_m)\theta(x_m)}{ma}
% 			 - \frac{\mu+\mu_{hc}}{m}\dot{x}_m  - f_d \text{sign}(\dot{x}_m)
% 			 - \frac{2\alpha U_p\ x_m}{mL}
% 			 - p_{hc,a}\ S_{hc} \\
% x_m = \dfrac{D_r(1-\cos(\theta))+2a\sin(\theta)}{2\cos(\theta)}\text{,}
% 				~~ \theta\leq\arctan\biggl(\dfrac{2a}{D_r}\biggr)\\
% \dfrac{U_p}{R_l} = \frac{2\alpha x_m \dot{x}_m}{L} - C_p\dot{U}_p 
% 	\end{dcases}
% 	\label{eq:OB+GPA+VH+piston}	
% % \end{small}
% \end{equation}
% \end{figure*}
% % \FloatBarrier
% %%%%%%%%%%%%%%%%%%%%%%%
    %///////////////////////////////////////////// 
	\subsection{Modeling of the hydraulic circuit}	
	\label{subsec:The hydraulic circuit}
    %/////////////////////////////////////////////

@@

The no leakage assumption imposes that all the fluid exiting the earplug is contained in the two hydraulic branches (eq. \ref{eq:conservation_masse}).
\begin{equation}
	q_{ear} = \dot{\Delta V_{ear}} = q_a + q_i
	\label{eq:conservation_masse}
\end{equation}
Moreover, the fluid entering the HC necessarily results on a piston displacement (eqs. \ref{eq:conservation_masse_ouvert},\ref{eq:conservation_masse_ferme}).
\begin{align}
	\text{Active cylinder ~}& \left\{~
	q_a = S_{hc} \dot{x}_m
	\right.
	\label{eq:conservation_masse_ouvert}\\
	\text{Inactive cylinder ~}& \left\{~
	q_i = S_{hc} \dot{x}_{ic}
	\right.
	\label{eq:conservation_masse_ferme}
\end{align}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
\section{NUMERICAL MODEL AND SIMULATIONS}
\label{sec:NUMERICAL MODEL AND SIMULATIONS}
%/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\/!\%
    %///////////////////////////////////////////// 
	\subsection{The harvester parameters setting}	
	\label{subsec:The harvester setting}
    %/////////////////////////////////////////////
The energy system combines many variable parameters. In order to determine them, it is first necessary to identify the fixed technological elements: 
\begin{itemize}
	\item The transducer is a APA50XS piezoelectric actuator, from Cedrat Technologies, exploited as a generator \cite{CEDRATTECHNOLOGIES2022}.
	\item The hydraulic actuation is ensured by the SMC MQP4-10S HCs which admit operates at 1kPa pressure \cite{SMC2022}.
	\item The usual hearing aid size is $50$x$20$x$10$mm \cite{Quattro2019}. The \mbox{$L=16mm$} parameter is chosen to be consistent with this scale. $\epsilon$ can still be adjusted with $x_0$.
\end{itemize}

@@

Figure \ref{fig:positions_Up_puissances_energies} additionally presents the different positions of the moving components, the APG voltage, the source and harvested powers and the different energies entering and exciting the system. The theoretical global efficiency of the harvester is evaluated at \mbox{$\eta_g=76$\%} with 20µJ harvested from one mastication. The energy losses are mostly located in the BR stage and depend on the quality factor $Q$ that has been set in accordance with previous work results on similar BR architectures \cite{Liu2013}. The energy source depends on the comfort pressure level $p_c$ that is set to 1kPa, i.e. a fraction of the theoretical maximal value of $12$kPa \cite{Bouchard-Roy2020}. The harvester theoretical efficiency could lead to a maximum of $684$µJ extractable from the ear canal.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={0cm 0cm 0cm 0.5cm},clip, width=\textwidth]{figures/positions+DeltaV_debits_pear.pdf}
	\caption{}
	\label{fig:positions+DeltaV_debits_pear}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[!htbp]
	\centering
	\captionsetup{justification=centering}
	\includegraphics[trim={0cm 0cm 0cm 0cm},clip, width=\textwidth]{figures/positions_Up_puissances_energies.pdf}
	\caption{}
	\label{fig:positions_Up_puissances_energies}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	
\begin{table}
\begin{subtable}[b]{.42\linewidth}
	\centering
	\resizebox{\linewidth}{!}{%
	\begin{tabular}{c|c}
\toprule
\multicolumn{1}{c}{~~\textbf{Parameter}~~}  & \multicolumn{1}{c}{~~\textbf{Value}~~} \\
\midrule
$K$	[N/µm]		&	0.256		\\	\hline
$x_0$ [mm]			&	0.49		\\	\hline
$L$	[mm]			&	16.0		\\	\hline
$m$	[g]				&	5.88		\\	\hline
$x_{c0}$ [mm]		&	0.69		\\	\hline
$R_l$ [k$\Omega$] 	&	6.39		\\	\hline
$\alpha$ [N/V]		&   0.105		\\	\hline
$C_p$ [µF]	   		&	0.25		\\	\hline
$Q$ [-]				&	50			\\	\hline
$\eta_g$ [\%]		& 	76			\\
\bottomrule
	\end{tabular}}
	\caption{Electromechanical}
	\label{tab:parametres électromécaniques}			
\end{subtable}
\hfillx
\begin{subtable}[b]{.55\linewidth}
	\centering
	\resizebox{\linewidth}{!}{%
	\begin{tabular}{c|c}
\toprule
\multicolumn{1}{c}{~~\textbf{Parameter}~~}  & \multicolumn{1}{c}{~~\textbf{Value}~~} \\
\midrule
$D_{hc}$ [mm] 	   			&	4						\\ \hline
$a_h$ [-]					&	29			 			\\ \hline
$p_c$ [kPa]					&	1				 		\\ \hline
$(\Delta V{ear})_{max}$ [mm$^3$]	& 60				\\ \hline
$Cf_0$ [Pa.s$^2$/m$^6$]			&	0.21e17			\\ \hline
$r_{Cf}$ [-]						&	26				\\ 
\bottomrule 
	\end{tabular}}
	\caption{Hydraulic}
	\label{tab:parametres_hydrauliques}		
\end{subtable}
\caption{Simulated model theoretical parameters}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

The device heart is the electromechanical converter composed of the BR with the APG. They have been designed at the application scale in order to prove their feasibility. The BR is fabricated by electrical discharge machining (EDM) in a monobloc structure with blades. This method optimizes the resonator quality factor by avoiding screw assemblies. Figure \ref{fig:BDT_OB+GPA} shows the BR mounted on the experimental test bench. The 8 hinges are assured by 4 buckled blades (BB) while a vertical guide blade (GB) ensures the APG mounting and avoids any of its rotations. The local thickenings at the midspan of the blades are intended to facilitate the fabrication. They also maximize the BB compression stiffness that maximizes harvester efficiency, as we show in the following section. The geometrical and mechanical parameters of the BR are referenced in table \ref{tab:parametres_lames}. The following sections describe the design method that led to the dimensions of the blades that depend on our needs.

@@

Soft hinge stiffness		   & $K_{\varphi}$             & 0.006 [N.m/rad] \\

@@

@@

The harvesting happens during the free oscillation phase. The main issue is to maximize the energy conversion efficiency $\eta_{ob}$ of this stage. Richards \emph{et al.} demonstrated that for the piezoelectric transduction at the resonance frequency, it can be approximated with the equation \ref{eq:eta_ob} \cite{Richards2004}. This expression is valid for an electric extraction based on the adaptive impedance matching strategy.

@@

$k^2_{sys}$ is the global electromechanical coupling coefficient of the harvesting system. The latter can be expressed with the electromechanical parameters of the APG and the stiffness of the blades (tab. \ref{tab:parametres_lames}) as follows :

@@

	k^2_{sys} = \dfrac{\alpha^2_{eq}}{\alpha^2_{eq} + (K+K_{gb})C_p}
	\label{eq:k2_sys}
Where $\alpha_{eq}$ is the equivalent piezoelectric force factor (PFF) given by equation \ref{eq:alpha_eq}. It depends on the APG PFF the and ratio between the equivalent stiffness $K_{eq}$ (eq. \ref{eq:K_eq}) of the structure on the $\vec{x}$ axis and the equivalent stiffness of the APG mounted on the GB.
	\alpha_{eq} = \alpha\ \dfrac{K_{eq}}{K + K_{gb}} 
	\label{eq:alpha_eq}
	K_{eq} = \dfrac{K_{bb}(K + K_{gb})}{K_{bb} + K_{gb} + K}
	\label{eq:K_eq}
Equations \ref{eq:eta_ob} - \ref{eq:K_eq} combined demonstrate that in order to maximize $\eta_{ob}$ we must make sure that $K_{bb} \gg K \gg K_{gb}$. If the condition is respected, then $K_{eq}$ can be assimilated to $K$.
